\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-08 paper/draft.tex
Finish 6.1

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of \ttt{width=}2--64.
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie with \ttt{eager} and \ttt{split} enabled, for \ttt{width=}\{2, 4, 8, 16, 32, 64\} are \{\todo{1.97, 3.5, 5.78, 8.83, 10.96, 13.47}\}$\times$, respectively.
For No-Eager, \ie for only \sys' transformations without its runtime support, speedups drop to 1.63, 2.54, 3.86, 5.93 7.46 9.35$\times$.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}
% we experiment with a $200\times$ level of parallelism to see potential overheads

Configurations where the parallel script produced is identical to an existing configuration are not shown.
There are two types of such cases (that appear in \todo{four} plots).
In the first, blue and red (not shown) lines are identical when \sys adds no \ttt{split} because the DFG width is constant for the entire pipeline;
  conversely, if both blue and red lines are shown, \ttt{split}s are needed due to changes in the pipeline width and result in performance improvements.
In the second type, the script features a command in \pur that introduces an aggregator resulting in width of 1 right after, needing a \ttt{split} to regain width.
Non-\ttt{split} configurations are sequential beyond that aggregator---\ie No Eager, Blocking Eager, and \sys w/o Split see no speedup and are identical (\todo{spell, bigrams}).

Tab.~\ref{tab:eval} shows that \sys's compilation time is negligible. % ($<5s$)
Filter and Bigrams-opt are not in Fig.~\ref{fig:microbenchmarks}, for which \sys achieves an average speedup 1.5--2.5$\times$ and $30\times$.
\sys achieves a COST~\cite{mcsherryscalability} of 2 for all benchmarks.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

@@

\heading{Discussion}
% TODO: Add
% * for scripts only with commands in \sta, we get linear speedup.
% * eager is better than no eager and blocking eager for all scripts
% * A \ttt{split} benefits scripts with \tpur or \tnpu commands, but does not negatively affect the rest.
% * \ttt{width}
% * why blocking is less than no eager
% ( standard vs non-standard Bigrams Fig.\ref{fig:microbenchmarks}).
\todo{These results explain a detail before getting the main points across: A is better than B, C is worse except G}
As discussed earlier (\S\ref{eval}, Parallelism), \sys does not have control over the input script's parallelism.
Thus, it often achieves the maximum possible speedup, \ie without unnecessary process spawns and context switches, earlier than the purported parallelism---$16-32\times$ for a 64-core system.
This is also due to the fact that \sys also adds its runtime primitives (\ttt{split} and \ttt{eager}), further increasing process spawns.
% \kk{eurosys: Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}
\todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}.
As described in Fig.\ref{optimizer}, these eager implementations perform worse than \sys's optimized \ttt{eager} relay in all cases (except for a negligible difference in Shortest-Scripts).

@@

Blocking eager: data parallelism, but no task parallelism
No eager: blocking FIFO does not allow pushing -> underutilize CPU

The reason why Blocking Eager is usually slower than No Eager is due to \ttt{agg}regators that read multiple input streams at the same time---such \ttt{sort -m}.

Sort-sort possibly shows the full spectrum:
  (iii) \sys's \ttt{eager} offers benefits despite the lack of \ttt{split} because it fully parallelizes the first \ttt{sort} while maintaining task parallelism
  (iv) \sys gets full benefits because \ttt{split}ting allows parallelizing the second \ttt{sort} too.

>>>>

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of \ttt{width=}2--64.
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie with \ttt{eager} and \ttt{split} enabled, for \ttt{width=}\{2, 4, 8, 16, 32, 64\} are \{\todo{1.97, 3.5, 5.78, 8.83, 10.96, 13.47}\}$\times$, respectively.
For \emph{No-Eager}, \ie \sys' transformations without its runtime support, speedups drop to 1.63, 2.54, 3.86, 5.93 7.46 9.35$\times$.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}
% we experiment with a $200\times$ level of parallelism to see potential overheads

Plots hide lines for configurations that lead to identical parallel programs.
There are two types of such cases (that appear in \todo{five} plots).
In the first type, \emph{Pash} (blue) and \emph{Pash-w/o-split} (red, hidden) lines are identical when \sys adds no \ttt{split} because the DFG width remains constant for the whole program;
  conversely, if both lines are shown, \ttt{split}s are needed due to changes in parallelism width and result in performance improvements.
In the second type, No Eager (green), Blocking Eager (orange), and \emph{Pash w/o Split} (red) are identical 
  because the program features a command in \pur relatively early.
This command requires an \ttt{agg}regator, whose output is of width 1, and beyond which non-\ttt{split} configurations are sequential (and see no speedup).
Conversely, all three lines exist when the width remains above 1 for the full program.
% needing a \ttt{split} to regain width.

Tab.~\ref{tab:eval} shows that \sys's transformation time is negligible, and its COST~\cite{mcsherryscalability} is 2. %  for all benchmarks.
Filter and Bigrams-opt, not in Fig.~\ref{fig:microbenchmarks}, achieve speedups of 1.5--2.5$\times$ and 2--30$\times$.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

@@

\heading{Discussion}
% TODO: Add
% * \ttt{width}
% * why blocking is less than no eager
As expected, scripts with commands only in \sta get linear speedup.
\sys's \ttt{split} benefits scripts with \tpur or \tnpu commands, without negative effects on the rest.
Its \ttt{eager} primitive improves over Eager and Blocking-eager for all scripts.
% Blocking eager: data parallelism, but no task parallelism; No eager: blocking FIFO does not allow pushing -> underutilize CPU
\emph{Blocking Eager} is usually slower than \emph{No Eager} due to \ttt{agg}regators that read multiple input streams at the same time---\eg \ttt{sort -m}.
Sort-sort illustrates the full spectrum of primitives:
  (iii) \sys's \ttt{eager} offers benefits despite the lack of \ttt{split} because it fully parallelizes the first \ttt{sort} while maintaining task parallelism
  (iv) \sys gets full benefits because \ttt{split}ting allows parallelizing the second \ttt{sort} too.

% \todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}.
% As described in Fig.\ref{optimizer}, these eager implementations perform worse than \sys's optimized \ttt{eager} relay in all cases (except for a negligible difference in Shortest-Scripts).

As described earlier, \sys often achieves the maximum possible speedup % \ie without unnecessary process spawns and context switches, 
  in \ttt{width} that is lower than the number of available cores---\ie \ttt{width=}16--32 for a 64-core system.
This is also because \sys's runtime primitives spawn new processes---\eg Sort with \ttt{width=}8 spawns 37 processes:
  8 \ttt{tr}, 8 \ttt{sort}, 7 \emph{aggregation}, and 14 \ttt{relay} processes.

\todo{standard vs non-standard Bigrams}

% \kk{eurosys: Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}

@@
